---
title: "For Next Donation"
author: "Shi Keqing"
output:
  html_document:
    code_folding: hide
    fig_width: 12
    highlight: tango
    number_sections: yes
    theme: sandstone
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Intro & Purpose
## Background
[DonorsChoose.org](https://www.donorschoose.org/) has funded over 1.1 million classroom requests through the support of 3 million donors, the majority of whom were making their *first-ever donation* to a public school. If DonorsChoose.org can motivate even a fraction of those donors to make *another donation*, that could have a huge impact on the number of classroom requests fulfilled.

## Mission
A good solution will enable DonorsChoose.org to build **targeted email campaigns recommending specific classroom requests to prior donors**. Part of the challenge is to assess the needs of the organization, uncover insights from the data available, and build the right solution for this problem.

\pagebreak

# Load Packages and Data Importing
## Packages to Use
For this mission, we need to library packages for data-manipulation, visualization, RMarkdown writing and  non-supervised algorithm.

```{r package loading, results='hide'}
# Reproducible-Research 
library('knitr')

# Data-Manipulation
library('data.table')
library('tibble')

# Visualization
library('ggplot2')
```

## Data Importing
All the data provided by csv format, in **6 files** with size of **3.71 GB** totally on Windows.

```{r data importing}
donations <- as.tibble(fread(input = "D:/DataSet/Kaggle/DonorsChoose/Donations.csv", header = TRUE, sep = ","))

donors <- as.tibble(fread("D:/DataSet/Kaggle/DonorsChoose/Donors.csv", header = TRUE, sep = ","))

projects <- as.tibble(fread("D:/DataSet/Kaggle/DonorsChoose/Projects.csv", header = TRUE, sep = ","))

resources <- as.tibble(fread("D:/DataSet/Kaggle/DonorsChoose/Resources.csv", header = TRUE, sep = ","))
                             
schools <- as.tibble(fread("D:/DataSet/Kaggle/DonorsChoose/Schools.csv", header = TRUE, sep = ","))

teachers <- as.tibble(fread("D:/DataSet/Kaggle/DonorsChoose/Teachers.csv", header = TRUE, sep = ","))
```

\pagebreak

# First Look of the Data Files{.tabset .tabset-fade .tabset-pills} 
## Project Data{.tabset}
This dataset contains information on all projects posted on [DonorsChoose.org](https://www.donorschoose.org/) site in the past five years and four months, between **2013-01-01** and **2018-05-01**. Each row in this dataset represents a single project and each column is an attribute of the project.

### structure 
```{r project data:glimpse}
glimpse(projects)
```

### summary
```{r project data:summary}
summary(projects)
```

### head
Belowing table neglects column **Project Essay** because of the column includes essay text for describing, not suitable and necessary for show here. 
```{r project data:head}
knitr::kable(head(projects[1:4]), format = "markdown")
knitr::kable(head(projects[c(5:6,8:9)]), format = "markdown")
knitr::kable(head(projects[10:15]), format = "markdown")
```

## Schools Data{.tabset}
This dataset contains information at a school level. Each row represents a single school. This dataset is joined with the project data using the **School ID** column.

### structure 
```{r schools data:glimpse}
glimpse(schools)
```

### summary
```{r schools data:summary}
summary(schools)
```

### head
```{r schools data:head}
knitr::kable(head(schools[1:4]), format = "markdown")
knitr::kable(head(schools[5:9]), format = "markdown")
```

## Teachers Data{.tabset}
This dataset contains information at a teacher level. Each row represents a single teacher. This dataset is joined with the project data using the **Teacher ID** column.

### structure 
```{r teachers data:glimpse}
glimpse(teachers)
```

### summary
```{r teachers data:summary}
summary(teachers)
```

### head
```{r teachers data:head}
knitr::kable(head(teachers), format = "markdown")
```

## Resources Data{.tabset}
For every project in the above dataset, there are one or more resources that is requested. This dataset contains the names of each resource in the project request and is joined with the dataset above using the **Project ID** column.

### structure 
```{r resources data:glimpse}
glimpse(resources)
```

### summary
```{r resources data:summary}
summary(resources)
```

### head
```{r resources data:head}
knitr::kable(head(resources), format = "markdown")
```


## Donations Data{.tabset}
For every project in the above dataset, there are one or more donations. This dataset contains each donation from a citizen donor and is joined with the dataset above using the **Project ID** column.

### structure 
```{r donations data:glimpse}
glimpse(donations)
```

### summary
```{r donations data:summary}
summary(donations)
```

### head
```{r donations data:head}
knitr::kable(head(donations[1:3]), format = "markdown")
knitr::kable(head(donations[4:7]), format = "markdown")
```

## Donors Data{.tabset}
This dataset contains information at a donor level. Each row represents a single donor. This dataset is joined with the donation data using the **Donor ID** column.

### structure 
```{r donors data:glimpse}
glimpse(donors)
```

### summary
```{r donors data:summary}
summary(donors)
```

### head
```{r donors data:head}
knitr::kable(head(donors), format = "markdown")
```

\pagebreak

# Missing Value

# Explore Data Analysis
Based on the files and tables above, we can see that there are a lot columns and infomations.So I do not think it's a good idea that we follow the traditional way to explore the data **one column by one column** for more infomations, not to mention we are really lack of domain knowledges and understanding of funding. In my opinion, focusing on some basic facts may discover more useful experiences, such as the following:
    + to be continued...

## Teacher Data EDA{.tabset}

### Teacher First Project Posted Date 
```{r Teacher First Project Posted Date}
teachers$Year_Month <- substr(teachers$`Teacher First Project Posted Date`, start = 1, stop = 7)
teachers_by_month <- as.data.frame(table(teachers$`Teacher Prefix`, teachers$Year_Month))
names(teachers_by_month) <- c("Prefix", "Year_Month", "Amount")
teachers_by_month <- teachers_by_month[which(teachers_by_month$Prefix != ""),]

month_break <- seq(as.Date("2002-09-01"), as.Date(Sys.Date()), by = "6 month")
teachers_by_month$Year_Month <- as.character(teachers_by_month$Year_Month)
teachers_by_month$Year_Month <- paste(teachers_by_month$Year_Month, "-01", sep = "")
teachers_by_month$Year_Month <- as.Date(teachers_by_month$Year_Month, "%Y-%m-%d")

ggplot(teachers_by_month, aes(x = `Year_Month`, y = `Amount`, colour = `Prefix`, group = `Prefix`)) + 
    geom_line(size = 1.0) + geom_point(size = 1.5) + scale_color_brewer(palette = "Set1") +
    scale_x_date(breaks = month_break) +
    theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1),
          legend.position = "top", 
          legend.text = element_text(colour = "lightblue4"),
          axis.text = element_text(colour = "royalblue3")) + 
    xlab("Year-Month") + ylab("Amount of Teacher First Project Posted")
```


```{r}
teachers$years <- (as.integer(strptime(Sys.Date(),"%Y-%m-%d")-
                                strptime(teachers$`Teacher First Project Posted Date`,"%Y-%m-%d")))/365
teachers_by_years <- as.data.frame(table(teachers$`Teacher Prefix`, round(teachers$years, digits = 3)))
names(teachers_by_years) <- c("Prefix", "Years of Expreiences", "Amount of Teachers")
teachers_by_years <- teachers_by_years[which(teachers_by_years$Prefix != ""),]

teachers_by_years$`Years of Expreiences` <- as.numeric(as.character(teachers_by_years$`Years of Expreiences`))

ggplot(teachers_by_years, aes(x = `Years of Expreiences`, fill = Prefix)) + geom_density(alpha = 0.3) +
    facet_grid(Prefix ~.)
    
```